Methods for Large Scale Learning to Rank

### 2014-03-10 report/Chapters/RelatedWork.tex
indeling ensemble learning related work klaar

\subsection{Parallelising Learning-to-Rank through ensemble learning}
Schapire proved in 1990 that a strong model can be generated by combining weak models through a procedure that he called boosting \cite{Schapire1990}. The invention of the boosting method resulted in an increasing focus in machine learning research on methods to combine multiple weak models, which are called ensemble learning methods. Well-known ensemble methods include boosting \cite{Schapire1990}, bagging \cite{Breiman1996}, AdaBoost \cite{Freund1997} and stacked generalisation (stacking) \cite{Wolpert1992}.\\

Ensemble methods can be used to parallelise learning methods by training the weak models in the ensemble on different nodes. In the Learning-to-Rank field parallelisation of model training through ensemble learning is frequently achieved using decision tree learners and often using the boosting ensemble method, jointly called \acf{GBDT}. \ac{GBDT}'s are shown to be able to achieve good accuracy in a Learning-to-Rank setting when used in a pairwise \cite{Zheng2007} or listwise \cite{Chen2008} setting.
\subsubsection{Stochastic Boosting}
Ye et al. \cite{Ye2009} described how to implement the \ac{SGBDT} process in a parallel manner using both MPI and Hadoop.  Experiments showed the Hadoop implementation to result into too expensive communication cost to be useful. Authors believed that these high communication costs were a result of the communication intensive implementation that was not well suited for the MapReduce paradigm. The MPI approach proved to be successful and obtained near linear speed-ups.\\

Kocsis et al. \cite{Kocsis2013} propose way to train multiple weak learners in parallel and extend those models that are likely to yield a good model when combined through boosting. Authors showed through theoretical analysis that the proposed algorithm asymptotically achieve equal performance to regular boosting. Using this parallel BoostingTree algorithm \ac{GBDT} models could be trained in parallel.\\

Tyree et al. \cite{Tyree2011} describe a way of parallelising \ac{GBDT} models for Learning-to-Rank where the boosting step is still executed sequentially, instead the construction of the regression trees themselves are parallelised. The parallel decision tree building is based on Ben-Haim and Yom-Tov's work on parallel construction of decision trees for classification \cite{Ben-Haim2010}. Decision trees are build layer-by-layer where the calculations needed for building each new layer in the tree are divided over the nodes using a master-worker paradigm. The data is partitioned of the workers, who compress their share into histograms and send these to the master. The master uses those histograms to approximate the split and build the next layer. The master than communicates this new layer to the workers who can use this new layer to compute new histograms. this process is repeated until the tree depth limit is reached. The tree construction algorithm parallelised with described master-worker method is \ac{CART} \cite{Breiman1984}. Speed-up experiments on the LETOR and the Yahoo! Learning to Rank challenge data sets were performed. This parallel \ac{CART}-tree building approach showed speed-up of up to 42x on shared memory machines and up to 25x on distributed memory machines.

\subsubsection{Stacked generalisation}
A \ac{DSN} is a processing architecture developed from the field of \emph{Deep Learning}, that uses the \ac{MSE} loss function that is easily optimisable. \ac{DSN} is based on the stacked generalization ensemble learning method \cite{Wolpert1992}, an approach where several learning models are stacked on top of each other in such a way that the output of a model is used as one of the input features of models higher up in the stack. Each layer in the stack learns has the task of learning the weights of the inputs of that layer. The close-form constraints between input and output weights allow the input weight matrices to be estimated in a parallel manner.

\subsubsection{Boosting wrapped in Bagging}
Pavlov et al. \cite{Pavlov2010} at Yandex developed a hybrid model that combines two methods from ensemble learning: bagging and boosting. Bagging \cite{Breiman1996}, also called bootstrap aggregating, is a ensemble learning method in which $m$ training sets $D_1..D_m$ are constructed from the training set $D$ by uniformly sampling data items from $D$ with replacement. The bagging method is parallelisable by training each $D_i$ with $i \in \{1..m\}$ on a different node. Boosting \cite{Friedman2002} is an ensemble learning method in which multiple weak learners are iteratively added together into one ensemble model in such a way that new models focus more on those data instances that were misclassified before. The boosting step itself is not parallelisable, but the authors state that learning a short boosted sequence on a single node is still a do-able task.\\

@@

>>>>

\subsection{Parallelising Learning-to-Rank through ensemble learning}
Schapire proved in 1990 that a strong model can be generated by combining weak models through a procedure that he called boosting \cite{Schapire1990}. The invention of the boosting method resulted in an increasing focus in machine learning research on methods to combine multiple weak models, which are called ensemble learning methods. Well-known ensemble methods include gradient boosting \cite{Friedman2001}, bagging \cite{Breiman1996}, AdaBoost \cite{Freund1997} and stacked generalisation (stacking) \cite{Wolpert1992}.\\

Ensemble methods can be used to parallelise learning methods by training the weak models in the ensemble on different nodes. In the Learning-to-Rank field parallelisation of model training through ensemble learning is frequently achieved using decision tree learners and often using the boosting ensemble method, jointly called \acf{GBDT}. \ac{GBDT}'s are shown to be able to achieve good accuracy in a Learning-to-Rank setting when used in a pairwise \cite{Zheng2007} or listwise \cite{Chen2008} setting.

\subsubsection{Gradient Boosting}
Gradient Boosting \cite{Friedman2001} is an ensemble learning method in which multiple weak learners are iteratively added together into one ensemble model in such a way that new models focus more on those data instances that were misclassified before.\\

Kocsis et al. \cite{Kocsis2013} propose way to train multiple weak learners in parallel by extending those models that are likely to yield a good model when combined through boosting. Authors showed through theoretical analysis that the proposed algorithm asymptotically achieve equal performance to regular gradient boosting. \ac{GBDT} models could be trained in parallel using their BoostingTree algorithm.\\

Tyree et al. \cite{Tyree2011} describe a way of parallelising \ac{GBDT} models for Learning-to-Rank where the boosting step is still executed sequentially, instead the construction of the regression trees themselves are parallelised. The parallel decision tree building is based on Ben-Haim and Yom-Tov's work on parallel construction of decision trees for classification \cite{Ben-Haim2010}. Decision trees are build layer-by-layer where the calculations needed for building each new layer in the tree are divided over the nodes using a master-worker paradigm. The data is partitioned of the workers, who compress their share into histograms and send these to the master. The master uses those histograms to approximate the split and build the next layer. The master than communicates this new layer to the workers who can use this new layer to compute new histograms. this process is repeated until the tree depth limit is reached. The tree construction algorithm parallelised with described master-worker method is \ac{CART} \cite{Breiman1984}. Speed-up experiments on the LETOR and the Yahoo! Learning to Rank challenge data sets were performed. This parallel \ac{CART}-tree building approach showed speed-up of up to 42x on shared memory machines and up to 25x on distributed memory machines.\\

Ye et al. \cite{Ye2009} described how to implement the stochastic \ac{GBDT} model in a parallel manner using both MPI and Hadoop. Stochastic \ac{GBDT} differs from regular \ac{GBDT} models by using stochastic gradient boosting instead of regular gradient boosting. Stochastic gradient boosting is a minor alteration to regular gradient boosting that, at each iteration of the algorithm, trains a base learner on a sub-sample of the training set that is drawn at random without replacement \cite{Friedman2002}. Experiments showed the Hadoop implementation to result into too expensive communication cost to be useful. Authors believed that these high communication costs were a result of the communication intensive implementation that was not well suited for the MapReduce paradigm. The MPI approach proved to be successful and obtained near linear speed-ups.

\subsubsection{Boosting wrapped in Bagging}
Some approaches combine both boosting and bagging. Bagging \cite{Breiman1996}, also called bootstrap aggregating, is a ensemble learning method in which $m$ training sets $D_1..D_m$ are constructed from the training set $D$ by uniformly sampling data items from $D$ with replacement. The bagging method is parallelisable by training each $D_i$ with $i \in \{1..m\}$ on a different node.\\

Pavlov et al. \cite{Pavlov2010} from Yandex were the first to propose a boosting-wrapped-in-bagging approach, which they called BagBoo. The boosting step itself is not parallelisable, but the authors state that learning a short boosted sequence on a single node is still a do-able task.

@@

\subsubsection{Stacked generalisation}
A \ac{DSN} is a processing architecture developed from the field of \emph{Deep Learning}, that uses the \ac{MSE} loss function that is easily optimisable. \ac{DSN} is based on the stacked generalization ensemble learning method \cite{Wolpert1992}, an approach where several learning models are stacked on top of each other in such a way that the output of a model is used as one of the input features of models higher up in the stack. Each layer in the stack learns has the task of learning the weights of the inputs of that layer. The close-form constraints between input and output weights allow the input weight matrices to be estimated in a parallel manner.\\
